---
title: "R Notebook"
output: html_notebook
---

# Monopoly Simulation

This is my attempt to find out which strategy for Monopoly is the best. The best way that I could think to do this was through a Monte Carlo simulation of the game. There may be other data sources of the game but this was an attempt to display my skills in coding and thinking. This is a simulation and while there may be shortcomings I feel that it is an accurate representation of the game. Some assumptions were made which I have documented. My work was made to follow the official rules of Monopoly as closely as possible. This file contains the data and simulation. Another file will be added with analysis of the data. While my code may be rudimentary at some points, for those of you looking at this code to see if I am fit for a position I would ask that you think more regarding my thought process in putting together the functions and and the complexity of the simulation. More comments and notes are provided throughout the code. 

Things outside of the scope of this project:

1) Bidding for properties.
2) Selling properties to other players when going bankrupt.
3) Mortgaging properties.

Assumptions made during simulation:

1) People leave jail their first time in there if they can afford it.
2) Number of Color Sets is only properties on which houses can be built.
3) Comfort levels are constant throughout the course of the game.
4) Players will only buy a property from another player if they already own a majority of the color set.
5) When buying properties from other players, they will buy from the player with the least amount of money.
6) Property markup values when player sells to another player.

```{r message = FALSE}
library(plyr)
library(dplyr)
library(tictoc)
library(future.apply)
plan(multisession, workers = 7)
```

## Player Setup

This function sets up three different aspects of the simulation. 
1) Properties Dataframe: This is the dataframe that most of the simulation takes place on. As the simulation runs adjustments are made to values within this dataframe. It contains:
 + Names: names of properties 
 + Ident: the property identifier (position on board) 
 + Status: the property status (1-7) 
 + Group: The group that the property belongs to (0-9)
 + Price: The purchase price
 + Rent: The rent when the property is owned
 + Rent_2: The rent when the entire color set is owned
 + Rent_1H - Rent_Ho: The rent with 1 - 4 house(s) and rent for 1 hotel
 + Mortgage: Mortgage Price
 + Owned.by: Which player the property is owned by (0 indicates no owner)
 + Running.rent: The rent charged when someone lands on that property
 + House.price: The price of building a house on the property
 + Type: The type of property indicates whether it is a property that can be built on or not
 + Money.made: The money made is a cumulative sum of the money made from that property each game
2) Other Properties: This is a data frame indicates what will be done when someone lands on "income tax", "jail", "luxury tax".
3) Player Setup: This player setup creates a series of vectors and lists in which information regarding players is saved throughout the game. Most importantly it sets up player comfort levels. Player comfort levels are what decisions for each player are based off. Each time a player lands on a property to be purchased, the decision to buy it is based off of the percentage of the player's cash on hand that purchase would be. For example, if a player one has $200 on hand, and the property is $50, then the decision to buy it is based off whether .25 is less than player one's comfort level. If it is, then then player one purchases the property, gets out of jail, buys the house, etc.

```{r}
setup <- function(){
  
#PROPERTY SETUP 
  names <- c("mediterranean avenue", "baltic avenue", "oriental avenue", " vermont avenue", "connecticut avenue", "st. charles place", "states avenue", "virginia avenue", "st. james place", "tennessee avenue", "new york avenue", "kentucky avenue", "indiana avenue", " illinois avenue", "atlantic avenue", "ventnor avenue", "marvin gardens", "pacific avenue", " north carolina avenue", "pennsylvania avenue", "park place", "boardwalk", "reading railroad", "pennsylvania railroad", "b&o railroad", "short line", "electric company", "water works")
  price <- c(60, 60, 100, 100, 120, 140, 140, 160, 180, 180, 200, 220, 220, 240, 260, 260, 280, 300, 300, 320, 350, 400, 200, 200, 200, 200, 150, 150)
  house_price <- c(50, 50, 50, 50, 50, 100, 100, 100, 100, 100, 100, 150, 150, 150, 150, 150, 150, 200, 200, 200, 200, 200, 0, 0, 0, 0, 0, 0)
  rent <- c(2, 4, 6, 6, 8, 10, 10, 12, 14, 14, 16, 18, 18, 20, 22, 22, 24, 26, 26, 28, 35, 50, 25, 25, 25, 25, 0, 0)
  rent_2 <- rent*2
  rent_1H <- c(rent[1:22] * 5, rep(100, 4), rep(0, 2))
  rent_2H <- c(30, 60, 90, 90, 100, 150, 150, 180, 200, 200, 220, 250, 250, 300, 330, 330, 360, 390, 390, 450, 500, 600, 200, 200, 200, 200, 0, 0)
  rent_3H <- c(90, 180, 270, 270, 300, 450, 450, 500, 550, 550, 600, 700, 700, 750, 800, 800, 850, 900, 900, 1000, 1100, 1400, 0, 0, 0, 0, 0, 0)
  rent_4H <- c(160, 320, 400, 400, 450, 625, 625, 700, 750, 750, 800, 875, 875, 925, 975, 975, 1025, 1100, 1100, 1200, 1300, 1700, 0, 0, 0, 0, 0, 0)
  rent_Ho <- c(250, 450, 550, 550, 600, 750, 750, 900, 950, 950, 1000, 1050, 1050, 1100, 1150, 1150, 1200, 1275, 1275, 1400, 1500, 2000, 0, 0, 0, 0, 0, 0)
  mortgage <- c(30, 30, 50, 50, 60, 70, 70, 80, 90, 90, 100, 110, 110, 120, 130, 130, 140, 150, 150, 160, 175, 200, 0, 0, 0, 0, 0, 0)
  ident <- c(1, 3, 6, 8, 9, 11, 13, 14, 16, 18, 19, 21, 23, 24, 26, 27, 29, 31, 32, 34, 37, 39, 5, 15, 25, 35, 12, 28)
  status <- rep(0, 28)
  owned.by <- rep(0, 28)
  running.rent <- rep(NA, 28)
  group <- c(1, 1, 2, 2, 2, 3, 3, 3, 4, 4, 4, 5, 5, 5, 6, 6, 6, 7, 7, 7, 8, 8, 9, 9, 9, 9, 0, 0)
  house.price <- c(rep(50, 5), rep(100, 6), rep(150, 6), rep(200, 5), rep(0, 6))
  type <- c(rep(0, 22), rep(1, 6))
  money.made <- c(rep(0, 28))
  
  properties <<- data.frame(names, ident, status, group, price, rent, rent_2, rent_1H, rent_2H, rent_3H, rent_4H, rent_Ho, mortgage, owned.by, running.rent, house.price, type, money.made)
  
#SETTING UP OTHER PROPERTIES
  Onames <- c("income tax", "jail", "luxury tax")
  Oident <- c(4, 10, 38)
  Ocost <- c(200, 0, 75)
  
  other_spots <<- data.frame(Onames, Oident, Ocost)

#PLAYER SETUP
  player.cash <<- c(1500, 1500, 1500, 1500, 1500, 1500)
  levels <- seq(0, 1, by = .01)
  player.comfort <<- sample(levels, 6)
  players <<- c(1:6)
  pos <<- rep(0, 6)
  turn <<- rep(0, 6)
  spot1 <- c()
  spot2 <- c()
  spot3 <- c()
  spot4 <- c()
  spot5 <- c()
  spot6 <- c()
  spots <<- list(spot1, spot2, spot3, spot4, spot5, spot6)
  bankrupt.players <<- rep(0, 6)
  total.value.fifteen <<- rep(0, 6)
  owned.blue <<- rep(0, 6)
  times.in.jail <<- rep(0, 6)
  properties.sold <<- rep(0, 6)
  color.sets <<- rep(0,6)
  props.at.five <<- rep(0,6)
  doubles <<- vector("list", 6)
  
#COMMUNITY CHEST SETUP
  chest.cards <<- c(1:17)
  get.out <<- c(0, 0)
  
#CHANCE SETUP
  chance.cards <<- c(1:16)
}
```

## Functions

These are the functions used throughout the simulation. 
```{r}
#The roll function simulates a roll of two die, adds them to spaces moved, and saves them.  
roll <- function(i){
  die1 <<- sample(6, 1)
  die2 <<- sample(6, 1)
  spaces <<- die1 + die2
  
  variable.costs(i)
  
  if (die1 == die2){
    doubles[[i]] <<- c(doubles[[i]], 1)
  } else {
    doubles[[i]] <<- c(doubles[[i]], 0)
  }
}

#The jail function runs a player through what happens when a player is in jail. A player is sent to jail when they roll doubles three times or lands on space 30. They are immediately sent to space 100 (a holder space to differentiate between when you are in jail and 'just visting'). A player will get out of jail on the first turn there if they can afford it, if they cannot then they will stay there until they roll doubles. 
jail <- function(i){
  if (tail(doubles[[i]], 1) == 1 || i %in% get.out) {
    
    pos[i] <<- 10
    
    space.action(i)
    
    property.check(i)
    
    rr.ut.status()
    
    buy.houses()
    
    rent()
    
  } else if (50 / player.cash[i] < player.comfort 
             || sum(spots[[i]][(max(length(spots[[i]]) - 2, 0)) : (length(spots[[i]]))] == 100) == 3) {
    
    player.cash[i] <<- player.cash[i] - 50
    
    bankrupt(i)
    
    pos[i] <<- 10
    
    if (bankrupt.players[i] == 0) {
      
      space.action(i)
      
      property.check(i)
      
      rr.ut.status()
      
      buy.houses()
      
      rent()
    }
    
  } else {
    
    turn[i] <<- turn[i] + 1
    
    spots[[i]] <<- c(spots[[i]], pos[i])
  }
  
}

#This function makes up a large portion of what happens on the board. It sets up they payout for passing 'go'. It sets up what happens if a player lands on a property that is unowned and gives the player the opportunity to purchase the property, if the property is purchased then the player pays rent to the owner. It also sets up what happens if a player lands on an 'other spot' such as luxury tax, income tax or 'go to jail'. 
space.action <- function(i){

  pos[i] <<- pos[i] + spaces
  
  if (pos[i] > 40) {
    pos[i] <<- pos[i] - 40
    player.cash[i] <<- player.cash[i] + 200
  }
  
  if(sum(tail(doubles[[i]], 3) != 3)){
    
    if (pos[i] %in% properties$ident) {
      a <<- which(properties$ident == pos[i])
      if (properties$status[a] >= 1) {
        player.cash[i] <<- player.cash[i] - properties$running.rent[a]
        bankrupt(i) 
        renter <<- properties$owned.by[a]
        
        if (player.cash[i] < 0){
          player.cash[renter] <<- player.cash[renter] + (properties$running.rent[a] + player.cash[i])
        } else {
          player.cash[renter] <<- player.cash[renter] + properties$running.rent[a]
        }
        
        properties$money.made[a] <<- properties$money.made[a] + properties$running.rent[a]
        
      } else if (properties$status[a] == 0) {
        if ((properties$price[a] / player.cash[i]) < player.comfort[i]) {
          player.cash[i] <<- player.cash[i] - properties$price[a]
          properties$status[a] <<- 1
          properties$owned.by[a] <<- i
        }  
      }
    } else if (pos[i] %in% other_spots$Oident) {
      b <<- which(other_spots$Oident == pos[i])
      player.cash[i] <<- player.cash[i] - other_spots$Ocost[b]
      bankrupt(i)
    } else if (pos[i] == 30) {
      pos[i] <<- 100
    } else if (pos[i] %in% c(2, 17, 33)){
      community.chest(i)
    } else if (pos[i] %in% c(7, 22, 36)){
      chance(i)
    }
  } else {
    pos[i] <<- 100
  }
  
  turn[i] <<- turn[i] + 1
 
  spots[[i]] <<- c(spots[[i]], pos[i])
}

#Property check function is the function that makes it so players can purchase properties from other players. Players will only buy a property if they own a majority or equal part of a color set. They will buy the properties from the player with the least amount of money on hand in the color set of interest. Properties are sold at a markup to other players. That markup is 3.5x's for buyers who have the same number of properties in that color set as the buyers, and 2.5x's for buyers who have more properties in that color set than the seller.
property.check <- function(i){
  if(i %in% properties$owned.by){
    #Also change it so that instead of checking against everyone elses properties you check against other peoples properties to see even if someone else owns two of them you could buy one from someone else
    prop.tab <- aggregate(status ~ group + owned.by, data = properties, sum)
    #remove the 0's
    prop.tab <- prop.tab[which(prop.tab$owned.by != 0), ]
    
    if(sum(prop.tab$owned.by != i) > 0){
      
      prop.for <- prop.tab[prop.tab$owned.by == i, ] 
      prop.against <- prop.tab[prop.tab$owned.by != i, ] 
      
      other.prop <- aggregate(status ~ group , data = prop.against, max)
      
      groups.to.buy <- c()
      
      for (y in prop.for$group) {
        if(y %in% other.prop$group && prop.for$status[which(prop.for$group == y)] >= other.prop$status[which(other.prop$group == y)]) {
          groups.to.buy <- c(groups.to.buy, y)
        }
      }
      
      if(length(groups.to.buy) > 0) {
        
        location <- c()
        
        for(j in 1:length(groups.to.buy)){
          location <- sort(c(which(properties$group == groups.to.buy[j]), location))
        }
        
        prop.to.buy <- properties$ident[location]
        sellers <- unique(properties$owned.by[location])
        sellers <- sellers[which(sellers != i)] 
        
        if (length(sellers) > 0) {
          poor <- which(min(player.cash[sellers]) == player.cash)
          buy.from <- intersect(sellers, poor)[1]
          
          purchase <- data.frame()
          
          for(c in 1:length(groups.to.buy)){
            purchase <- rbind(properties %>% filter(group == groups.to.buy[c] & owned.by == buy.from), purchase)
          }
          
          purchase <- purchase[sample(length(purchase$ident), 1),]
          
          g2b <- purchase$group
          
          if(other.prop$status[which(other.prop$group == g2b)] == prop.for$status[which(prop.for$group == g2b)]){
            markup <- 3.5
          } else {
            markup <- 2.5
          }
          
          player.cash[i] <<- player.cash[i] - (purchase$price * markup) 
          player.cash[buy.from] <<- player.cash[buy.from] + (purchase$price * markup)
          properties.sold[buy.from] <<- properties.sold[buy.from] + 1
          trans <- which(properties$ident == purchase$ident)
          properties$owned.by[trans] <<- i
        }
      }
    }
  }
}

#The rent function just assigns running rent for each property based on the property status.
rent <- function(){
  for (c in 1:length(properties$rent)){
    if (properties$status[c] == 1) {
      properties$running.rent[c] <<- properties$rent[c] 
    } else if (properties$status[c] == 2) {
      properties$running.rent[c] <<- properties$rent_2[c]
    } else if (properties$status[c] == 3) {
      properties$running.rent[c] <<- properties$rent_1H[c]
    } else if (properties$status[c] == 4) {
      properties$running.rent[c] <<- properties$rent_2H[c]
    } else if (properties$status[c] == 5) {
      properties$running.rent[c] <<- properties$rent_3H[c]
    } else if (properties$status[c] == 6) {
      properties$running.rent[c] <<- properties$rent_4H[c]
    } else if (properties$status[c] == 7) {
      properties$running.rent[c] <<- properties$rent_Ho[c]
    } else {
      properties$running.rent[c] <<- 0
    }
  }
}

#Buy.houses funciton is the process through which players buy houeses. This can be done at any point regardless of who's turn it is. Playes will buy houses on full color sets if they can afford it and given their comfort level. Players must build evenly, this is an added challenge because it means that you cannot build a property all the way up to a hotel if the property next to it doesn't even have a hosue. This fucntion assigns a status of 2 to all full color sets and increases the status to 3, 4, etc. when a houses is added. 
buy.houses <- function(){
  
  count.buildings()
  
  for (c in 1:8) {
    prop.grab <- which(properties$group == c)
    
    #moving to status 2... aka a full color set
    if (length(unique(properties$owned.by[prop.grab])) == 1 && properties$status[prop.grab] == 1){
      properties$status[prop.grab] <<- 2
      color.sets[unique(properties$owned.by[prop.grab])] <<- color.sets[unique(properties$owned.by[prop.grab])] + 1
    }
      
      prop.up <- properties[prop.grab, ]
      candidates <- prop.up[which(properties$status[prop.grab] == min(properties$status[prop.grab])), ]
      upgrade <- candidates[1, ]
      change <- which(properties$ident == upgrade$ident)
      p <- properties$owned.by[change]
      
      if (between(upgrade$status, 2, 6) && (player.cash[p] * player.comfort[p]) > properties$house.price[change]) {
        if(properties$status[change] == 6 && hotel.count < 12){
          properties$status[change] <<-  properties$status[change] + 1
          player.cash[p] <<- player.cash[p] - properties$house.price[change]
        } else if(properties$status[change] < 6 && house.count < 32){
          properties$status[change] <<-  properties$status[change] + 1
          player.cash[p] <<- player.cash[p] - properties$house.price[change]
        }
      }
  }
}

#The rr.ut.status function stands for railroad and utilities. Status for railroads and utilities do not work like status for properties on which houses can be built. These properties move up in status based on how many of them you own. For example when you own two railroads your status the status of those railroads goes up to 2 eventhough you don't own all of the railroads. This increases the rent. This assigns statuses to each property based on how many different players own them.
rr.ut.status <- function(){
  rrs <- which(properties$group == 9)
  for (c in rrs){
    if (properties$status[c] != 0){
      properties$status[c] <<- sum(properties$owned.by[c] == properties$owned.by[rrs])
    }
  }
  
  utilities <- which(properties$group == 0)
  for (c in utilities){
    if (properties$status[c] != 0){
      properties$status[c] <<- sum(properties$owned.by[c] == properties$owned.by[utilities]) 
    }
  }
}

#The sell.houses function is part of the larger bankrupt function. When a player lands on a place in which he will have to pay but has insufficient funds they begin to sell their properties. The sell.houses function sells hotels and houses evenly within a color group beginning from the cheapest color group that that player owns. Hotels and houses are sold within that color group until the player can either pay the bill or all statuses within that color group are 2. Note: The rent function at the end to re-assign running.rent values.
sell.houses <- function(i){
  first <- which(properties$owned.by == i)
  second <- which(properties$type == 0)
  while(player.cash[i] < 0 && sum(c(3, 4, 5, 6, 7) %in% properties$status[intersect(first, second)]) >= 1){
    to.sell <- properties[which(properties$owned.by == i),] 
    to.sell <- to.sell[intersect(which(to.sell$status >= 3), which(to.sell$type == 0)),]
    group.grab <- min(to.sell$status)
    to.sell <- to.sell[which(to.sell$group %in% to.sell$group[which(to.sell$status == group.grab)]), ]
    to.sell <- to.sell[which(to.sell$house.price == min(to.sell$house.price)),]
    to.sell <- to.sell[which(to.sell$group == to.sell$group[1]),]
    
    #sell one property
    downgrade <- to.sell[which(to.sell$status == max(to.sell$status))[1], ]
    change <- which(to.sell$ident == downgrade$ident)
    to.sell$status[change] <- to.sell$status[change] - 1
    
    change <- which(properties$ident == downgrade$ident)
    properties$status[change] <<- properties$status[change] - 1
    
    player.cash[i] <<- player.cash[i] + (properties$house.price[change] / 2)
    
    while(player.cash[i] < 0 && min(to.sell$status) != max(to.sell$status)){
      highest <- to.sell[to.sell$status == max(to.sell$status), ]
      
      downgrade <- highest[1, ]
      change <- which(to.sell$ident == downgrade$ident)
      to.sell$status[change] <- to.sell$status[change] - 1
      
      change <- which(properties$ident == downgrade$ident)
      properties$status[change] <<- properties$status[change] - 1
      
      player.cash[i] <<- player.cash[i] + (properties$house.price[change] / 2)
    }
    rent()
  }
}

#The sell.ones function is also part of the bankrupt function. This function will sell of the properties that player owns of status one and assigns them a status of 0, which can then be purchased again by other players. Note: The rent function at the end to re-assign running.rent values.
sell.ones <- function(i){
  if(length(intersect(which(properties$status == 1), which(properties$owned.by == i))) > 0) {
    ones <- properties[intersect(which(properties$status == 1), which(properties$owned.by == i)),]
    for (c in 1:length(ones$names)){
      if(player.cash[i] < 0){
        change <- which(properties$ident == ones$ident[c])
        player.cash[i] <<- player.cash[i] + properties$price[change]/2 
        properties.sold[i] <<- properties.sold[i] + 1 
        properties$status[change] <<- 0 
        properties$owned.by[change] <<- 0
      }
    }
    rent()
  }
}

#The two.to.one function sells one property in a full color set, making it's status 0, and assigns the rest of that color set a status of 1, which can then be sold in the sell.ones function. This only takes place while cash is less than 0. It is also part of the larger bankrupt function. Note: The rent function at the end to re-assign running.rent values.
two.to.one <- function(i){
  color.s <- properties[intersect(which(properties$owned.by == i), which(properties$status == 2)),]
  for(g in unique(color.s$group)){
    if(player.cash[i] < 0){
      color.sell <- color.s[which(color.s$group == g),]
      
      #sell one of them to the bank
      change <- which(properties$ident == color.sell$ident[1])
      player.cash[i] <<- player.cash[i] + (properties$price[change] / 2) #make these <<-
      properties.sold[i] <<- properties.sold[i] + 1
      properties$status[change] <<- 0 #make these <<-
      properties$owned.by[change] <<- 0
      color.sell$status[1] <- 0
      
      #all other properties become 1's sell these while player.cash < 0
      for (c in 2:length(color.sell$status)){
        change <- which(color.sell$ident[c] == properties$ident)
        properties$status[change] <<- 1
      }
      
      sell.ones(i)
      
      rent()
    }
  }
}

#The other.to.one function is part of the larger bankrupt function. It sells railroad and utilities properties and updates their status while a player is bankrupt. It does this until the player's cash is aboce 0.
other.to.one <- function(i) {
  if (player.cash[i] < 0){
    for (c in c(9, 0)){
      spots <- which(properties$group == c)
      while(player.cash[i] < 0 && sum(properties$owned.by[spots] == i) > 0) {
        to.sell <- intersect(which(properties$owned.by == i), which(properties$group == 9))[1]
        player.cash[i] <<- player.cash[i] + (properties$price[to.sell] / 2) # make these <<-
        properties.sold[i] <<- properties.sold[i] + 1
        properties$status[to.sell] <<- 0
        properties$owned.by[to.sell] <<- 0
      }
    }
    rr.ut.status()
  }
}

#The bankrupt function is a combination of the sell.ones, sell.houses, two.to.one, other.to.one, rent, and rr.ut.status functions. It is the process through which players systematically sell their properties so that they can pay a bill when they don't have the money to do so. If, after they have sold all of their properties, they still have a negative amount of cash, then the player is labeled as bankrupt and removed from the game. The winner is never labeled as bankrupt and remains in the bankrupt.players vector as 0. The rest of the players are labeled in order that they went bankrupt (ie. the player with the value 1 was bankrupt first and the player with value 5 was the last to go bankrupt.)
bankrupt <- function(i){
  if (player.cash[i] < 0) {
    
    sell.ones(i)
    
    sell.houses(i)
    
    two.to.one(i)
    
    other.to.one(i)
    
    sell.ones(i)
    
    if(player.cash[i] < 0){
        bankrupt.players[i] <<- sum(bankrupt.players > 0) + 1
      }
    
  }
}

#The property.costs function creates a vector called prop.value which is a running total of the total values of each space on the board. It is the sum of the price of the property times the number of houses/hotel built on it. IT is used to calculate a player's total value. It is rolled into the variable costs function, while not a variable cost it is assigned at the same time becasue income tax is calculated based off of the property's total value.
property.costs <- function(i){
  new.status <- c()
  for (i in 1:length(properties$status)) {
    if (properties$status[i] < 3) {
      new.status[i] <- 0
    } else {
      new.status[i] <- properties$status[i] - 2
    }
  }
  
  prop.value <<- properties$price + (new.status * properties$house.price)
}

#The variable.costs function assigns variable costs to utilities based on the value of the two dice as outlined in the monopoly rules. It then calculates the total value of the properties owned by whoever's turn it is and uses that to make a decision for whether or not the player will chose to pay the $200 penalty or 10% penalty when the player lands on income tax.
variable.costs <- function(i) {
  #public works variable costs
  properties$rent[27:28] <<- spaces * 4
  properties$rent_2[27:28] <<- spaces * 10
  
  property.costs()
  
  #income tax variable costs
  p <- which(properties$owned.by == i)
  ten.perc <- .1 * sum(prop.value[p])
  
  if (ten.perc < 200){
    other_spots$Ocost[1] <<- ten.perc
  } else {
    other_spots$Ocost[1] <<- 200
  }
  
}

#This is the community chest function. It randomly selects from the cards available that have not been selected before then it goes through all of the different possible cards with their assigned penalties. Some just affect the player. Some affect all the players. The most tricky one to code penalized players for all of their houses and hotes.
community.chest <- function(i){
  
  if(length(chest.cards) == 0){
    chest.cards <<- c(1:17)
  }
  
  draw <- sample(chest.cards, 1)
  chest.cards <<- chest.cards[-draw]
  
  if(draw == 1){
    pos[i] <<- 40
  } else if (draw == 2){
    player.cash[i] <<- player.cash[i] + 200
  } else if (draw == 3){
    player.cash[i] <<- player.cash[i] + 50
  } else if (draw == 4){
    player.cash <<- player.cash - 50
    player.cash[i] <<- player.cash[i] + 100
    bankrupt(i)
  } else if (draw %in% c(5, 6, 7)){
    player.cash[i] <<- player.cash[i] + 100
  } else if (draw == 8){
    player.cash[i] <<- player.cash[i] + 20
  } else if (draw == 9){
    player.cash <<- player.cash - 10
    player.cash[i] <<- player.cash[i] + 20
    bankrupt(i)
  } else if (draw == 10){
    player.cash[i] <<- player.cash[i] + 25
  } else if (draw == 11){
    player.cash[i] <<- player.cash[i] + 10
  } else if (draw %in% c(12, 13, 14)){
    player.cash[i] <<- player.cash[i] - 50 
    bankrupt(i)
  } else if (draw == 15){
    hos.and.hot <- properties$status[which(properties$owned.by == i)]
    
    if(length(hos.and.hot) > 0){
      
      hotels <- sum(hos.and.hot == 7)
      houses <- sum(hos.and.hot[which(between(hos.and.hot, 3, 6))] - 2)
      
      payment <- hotels * 115 + houses * 40
      
      player.cash[i] <<- player.cash[i] - payment
      
      bankrupt(i)
    }
  } else if (draw == 16){
    pos[i] <<- 100
  } else if (draw == 17){
    get.out[1] <<- i
  }

}

# This function is for a piece of code that is run repeatedly in the chance function. It advances players to the next railroad and lets them buy it if it is unowned or charges them twice the normal rent if it is owned.
rr.chance <- function(x, ij = i){
  pos[ij] <<- x
  rr <- which(properties$ident == x)
  if(properties$status[rr] > 0 ){
    cost <- 2 * (properties$running.rent[rr])
    player.cash[ij] <<- player.cash[ij] - cost
    player.cash[properties$owned.by[rr]] <<- player.cash[properties$owned.by[rr]] + cost
    bankrupt(ij)
  } else {
    if ((properties$price[rr] / player.cash[ij]) < player.comfort[ij]) {
      player.cash[ij] <<- player.cash[ij] - properties$price[rr]
      properties$status[rr] <<- 1
      properties$owned.by[rr] <<- ij
      bankrupt(ij)
    }  
  }
}

# This function is for a piece of code that is run repeatedly in the chance function. It advances players to the next utility and lets them buy it if it is unowned or ten times the roll of the die if it is owned.
util.chance <- function(x, ij = i){
  pos[ij] <<- x
  ut <- which(properties$ident == x)
  if(properties$status[ut] > 0 ){
    cost <- 10 * (sample(1:6, 1) + sample(1:6, 1))
    player.cash[ij] <<- player.cash[ij] - cost
    player.cash[properties$owned.by[ut]] <<- player.cash[properties$owned.by[ut]] + cost
    bankrupt(ij)
  } else {
    if ((properties$price[ut] / player.cash[ij]) < player.comfort[ij]) {
      player.cash[ij] <<- player.cash[ij] - properties$price[ut]
      properties$status[ut] <<- 1
      properties$owned.by[ut] <<- ij
      bankrupt(ij)
    }  
  }
}

# The chance function is very similar to the community chest function. It simulates a draw from the chance stack and runs through the different chance cards.
chance <- function(i){
  
  if(length(chance.cards) == 0){
    chance.cards <<- c(1:16)
  }
  
  draw <<- sample(chance.cards, 1)
  chance.cards <<- chance.cards[-draw]
  
  if(draw == 1){
    pos[i] <<- 40
  } else if (draw == 2){
    if(pos[i] > 24){
      spaces <<- 24 + (40 - pos[i])
      space.action(i)
    } else {
      spaces <<- 24 - pos[i]
      space.action(i)
    }
  } else if (draw == 3){
    if(pos[i] > 11){
      spaces <<- 11 + (40 - pos[i])
      space.action(i)
    } else {
      spaces <<- 11 - pos[i]
      space.action(i)
    }
  } else if (draw == 4){
    if(between(pos[i], 13, 27)){
      util.chance(x = 28)
    } else {
      util.chance(x = 12)
    }
  } else if (draw == 5){
    if (between(pos[i], 5, 14)){
      rr.chance(x = 15) 
    } else if(between(pos[i], 15, 24)){
      rr.chance(x = 25)
    } else if(between(pos[i], 25, 34)){
      rr.chance(x = 35)
    } else {
      rr.chance(x = 5) 
    }
  } else if (draw == 6){
    spaces <<- -3
    space.action(i)
  } else if (draw == 7){
    player.cash[i] <<- player.cash[i] + 50
  } else if (draw == 8){
    get.out[2] <<- i
  } else if (draw == 9){
    pos[i] <<- 100
  } else if (draw == 10){
    hos.and.hot <- properties$status[which(properties$owned.by == i)]
    
    if(length(hos.and.hot) > 0){
      
      hotels <- sum(hos.and.hot == 7)
      houses <- sum(hos.and.hot[which(between(hos.and.hot, 3, 6))] - 2)
      
      payment <- hotels * 100 + houses * 25
      
      player.cash[i] <<- player.cash[i] - payment
      
      bankrupt(i)
    }
  } else if (draw == 11){
    player.cash[i] <<- player.cash[i] - 15
  } else if (draw == 12){
    if(pos[i] > 5){
      spaces <<- 40 - pos[i] + 5
      space.action(i)
    } else {
      spaces <<- 5 - pos[i]
      space.action(i)
    }
  } else if (draw == 13){
    spaces <<- 39 - pos[i]
    space.action(i)
  } else if (draw == 14){
    player.cash <<- player.cash + 50
    player.cash[i] <<- player.cash[i] - (50 * length(players))
    bankrupt(i)
  } else if (draw == 15){
    player.cash[i] <<- player.cash[i] + 150
  } else if (draw == 16){
    player.cash[i] <<- player.cash[i] + 100
  }
}

# This count buildings function is used to make sure that there are not more
# hotels and houses sold than the game will allow for. This function counts
# the number of houses and hotels in play so that a constraint can be set.
count.buildings <- function(){
  hotel.count <<- length(which(properties$status > 6))
  house.count <<- sum(properties$status[which(between(properties$status, 3, 6))])
}

# The Stale function is implemented because of the simple strategy of the 
# players it is possible for the players to reach a stale mate.The solution
# that I have come up with that is within the scope of the project is to
# set a cap on the number of turns a game can go to (500) and then to 
# calculate the net worth of the players and assigns the winners and 
# runners up from there.
stale <- function(){
  if(sum(bankrupt.players > 0) < length(players) - 1){
    rem.players <- which(bankrupt.players == 0)
    
    net.worth <<- c()
    for(j in rem.players){
      net.worth <<- c(net.worth, sum(prop.value[which(properties$owned.by == j)], player.cash[j]))
    }
    
    tie.breaker <- data.frame(cbind(rem.players, net.worth))
    tie.breaker <- tie.breaker[order(-net.worth), ]
    tie.breaker <- cbind(tie.breaker, "ident" = c(1:nrow(tie.breaker)) - 1)
    tie.breaker <- tie.breaker[-1, ]
    tie.breaker$ident <- 6 - tie.breaker$ident
    
    for(j in 1:nrow(tie.breaker)){
      bankrupt.players[tie.breaker$rem.players[j]] <<- tie.breaker$ident[j]
    }
    stale.mate <<- rep(1, 6)
  }
}
```

## Game Simulation

With all of the functions loaded now, it is time to run the monopoly simulation. For ease, the monopoly simulation is rolled into a monopoly function. The function can be run once, or in a Monte Carlo simulation as seen below. The simulation runs while there are less than 5 players bankrupt, simulating each players turn in a 'for' loop. Once a player is bankrupt, he is no longer run through the loop. At the end metrics are gathered on each of the players and properties. Two dataframes are output at the end of the function. You will see three interactions of the code, this is necessary because when players roll doubles they go again, and if they roll doubles three times in a row they are sent to jail.

The prop.df dataframe returns the number of times the property was landed on, whether the property was owned by the winner at the end of the game, the amount of money the property made throughout the game, and the group that the property belongs to.

The player.df dataframe returns the player, the cash held at the end of the game, the comfort level the player played with (assigned randomly at intervals of 0.01), the order the player went bankrupt in (remember: 0 is the winner), the number of turns taken, the players total value on the 15th turn, whether or not the player owned 'Park Place' or 'Boardwalk' throughout the course of the game, number of times in jail, number of properties sold, number of full color sets owned, and number of properties owned on the 5th turn. These metrics are useful in further analysis that will be performed.

```{r}
monopoly <- function() {
  
  setup()
  
  while(sum(bankrupt.players > 0) < 5 && sum(turn > 500) == 0) {
    
    rent()
    
    for (i in players) {
      i <<- i
      if (bankrupt.players[i] == 0) {
        
        roll(i)
        
        if (pos[i] == 100) {
          
          jail(i)
          
        } else {
          
          space.action(i)
          
          property.check(i)
          
          rr.ut.status()
          
          buy.houses()
          
          rent()
          
        }
        
        if(sum(tail(doubles[[i]], 1)) == 1 && bankrupt.players[i] == 0 
           && sum(tail(spots[[i]], 2) == 100) == 0) {
          
          while(tail(doubles[[i]], 1) == 1 && pos[i] != 100){
            
            roll(i)
            
            space.action(i)
            
            property.check(i)
            
            rr.ut.status()
            
            buy.houses()
            
            rent()
            
          }
        }
        
        
        
        if (turn[i] == 14) {
          total.value.fifteen[i] <- sum(prop.value[which(properties$owned.by == i)], player.cash[i])
        }
        
        if (sum(properties$owned.by[which(properties$group == 8)] > 0) > 0) {
          owned.blue[unique(properties$owned.by[which(properties$group == 8)])] <- 1
        }
        
        if (turn[i] == 4) {
          props.at.five[i] <- sum(properties$owned.by == i)
        }
      }
    }
  }
  
  stale.mate <<- rep(0, 6)
    
  stale()
  
  for (i in players) {
    times.in.jail[i] <- sum(spots[[i]] == 100)
  }
  
  prop.order <- properties[order(properties$ident),]
 
  times.landed.on <- as.data.frame(table(unlist(spots)))
  times.landed.on$Var1 <- as.numeric(times.landed.on$Var1)
  
  times.landed.on <- times.landed.on[intersect(prop.order$ident, times.landed.on$Var1), ]
  
  if(length(times.landed.on$Var1) != length(prop.order$ident)){
    for (c in prop.order$ident){
      if((c %in% times.landed.on[,1]) == FALSE) {
        times.landed.on <- rbind(times.landed.on, c(c, 0))
        times.landed.on <- times.landed.on[order(times.landed.on$Var1), ]
      }
    }
  }
  
  winner <- which(bankrupt.players == 0)
  winning.prop <- prop.order$owned.by == winner
  
  prop.df <<- cbind(times.landed.on, winning.prop, prop.order$money.made, prop.order$group)
  
  results.df <<- data.frame(players, player.cash, player.comfort, bankrupt.players, turn, 
                           total.value.fifteen, owned.blue, times.in.jail, properties.sold, color.sets, props.at.five, stale.mate)
  
  return(list(prop.df, results.df))
}
```

## One Game Simulation
This is the simulation of one game in monopoly. It will return two dataframes with results from the game.
```{r}
#monopoly()
```


## Monte Carlo Simulation

This code runs a Monte Carlo simulation study and compiles all results into two dataframes and writes them as '.csv' documents. It is commented out so that you can run the full R-Markdown file without the 1.5 hour long simulation.
```{r}
tic()
n.reps <- 3000
monte.carlo <- future_replicate(n.reps, monopoly())

prop.results.df <- data.frame()
player.results.df <- data.frame()

for (n in 1:n.reps) {
  prop.results.df <- rbind(prop.results.df, monte.carlo[[1, n]])
}

for (n in 1:n.reps){
  player.results.df <- rbind(player.results.df, monte.carlo[[2, n]])
}
toc()

write.csv(player.results.df, "sim.player.csv")
write.csv(prop.results.df, "sim.property.csv")
```
