---
title: "Monopoly Analysis"
output: html_notebook
---

# THIS PROJECT IS UNFINISHED AND IS NOT AN ACCURATE REPRESENTATION OF THE LEVEL OF WORK THAT I PRODUCE

```{r}
library(ggplot2)
```

```{r}
player <- read.csv('sim.player.up.csv', header = TRUE, stringsAsFactors = FALSE)[, 2:12]
property <- read.csv('sim.property.up.csv', header = TRUE, stringsAsFactors = FALSE)[, 2:6]

names(property) <- c("ident", "freq", "winning", "money.made", "group")
```

Which property is landed on the most?
```{r}
avg.landed <- aggregate(freq ~ ident, data = property, mean)
avg.landed <- avg.landed[order(avg.landed$freq, decreasing = TRUE), ]
head(avg.landed)
```

Is one property landed on significantly more than another?
```{r}
lm.prop.freq <- lm(freq ~ ident, data = property)
anova(lm.prop.freq)
```

Which group is landed on the most?
```{r}
group.freq <- aggregate(freq ~ group, property, mean)
group.freq[order(group.freq$freq, decreasing = TRUE), ]
```

Is one group landed on significantly more than another?
```{r}
lm.freq <- lm(freq ~ group, data = property)
anova(lm.freq)
```

Which property makes the most money?
```{r}
avg.money <- aggregate(money.made ~ ident, data = property, mean)
avg.money <- avg.money[order(avg.money$money.made, decreasing = TRUE), ]
head(avg.money)
```

Does one property make signficantly more money than another?
```{r}
prop.profit.lm <- lm(money.made ~ ident, data = property)
anova(prop.profit.lm)
```

Which group is the most profitable?
```{r}
group.money <- aggregate(money.made ~ group, data = property, mean)
group.money <- group.money[order(group.money$money.made, decreasing = TRUE), ]
head(group.money)
```

Is one group part of the winning players ownership more than another?
```{r}
lm.group <- lm(winning ~ group, data = property)
anova(lm.group)
```

Data Manipulation
```{r}
player$winner <- player$bankrupt.players == 0
winners <- player[which(player$winner == TRUE), ]
losers <- player[which(player$winner != TRUE), ]
```

What is the best player comfort level?
```{r}
ggplot(data = winners, mapping = aes(x = player.comfort)) +
  geom_histogram(binwidth = 0.05) +
  xlab("Comfort Level") +
  ylab("Count") +
  ggtitle("Player Comfort Histogram")
```

Does player order matter?
```{r}
table(winners$players)

table(winners$players) / length(winners$players)
```

Do winners own boardwalk or parkplace more often than the losers?
```{r}
mean(winners$owned.blue)

mean(losers$owned.blue)
```

Were winners in jail less?
```{r}
mean(winners$times.in.jail)

mean(losers$times.in.jail)
```

Average cash on hand at the end of the game?
```{r}
mean(winners$player.cash)
```

Average Number of Turns per game of the winner?
```{r}
mean(winners$turn)
```

